---
title: "JHU_Rm Quiz 4"
author: "mbh"
date: "Tuesday, June 23, 2015"
output: html_document
---

## Question One
Consider the space shuttle data ?shuttle in the MASS library. Consider modeling the use of the autolander as the outcome #(variable name use). Fit a logistic regression model with autolander (variable auto) use (labeled as "auto" 1) versus not (0) as predicted by wind sign (variable wind). Give the estimated odds ratio for autolander use comparing head winds, labeled as "head" in the variable headwind (numerator) to tail winds (denominator).

```{r q1 shuttle}
library(MASS)
load(shuttle)
str(shuttle)
shuttle$auto <- as.numeric(shuttle$use=="auto")
model1 <- glm(auto ~ wind-1,  family=binomial,  data=shuttle)
summary(model1)
exp(model1$coef[1]-model1$coef[2])
```



##Question Two

Consider the previous problem. Give the estimated odds ratio for autolander use comparing head winds (numerator) to tail winds (denominator) adjusting for wind strength from the variable magn.

```{r q2 shuttle}
shuttle$magn <- as.numeric(shuttle$magn)
model2<-glm(auto~wind+magn-1,family="binomial",data=shuttle)
summary(model2)
exp(model2$coef[1]-model2$coef[2])
```

##Question Three

If you fit a logistic regression model to a binary variable, for example use of the autolander,
then fit a logistic regression model for one minus the outcome (not using the autolander) 
what happens to the coefficients?

```{r q3 shuttle}
model1 <- glm(auto ~ wind,  family=binomial,  data=shuttle)
summary(model1)
shuttle$noauto<- as.numeric(shuttle$use=="noauto")
model3<-glm(noauto~ wind,family="binomial",data=shuttle)
summary(model3)
```

##Question Four

Consider the insect spray data InsectSprays. Fit a Poisson model using spray as a factor level. Report the estimated relative rate comapring spray A (numerator) to spray B (denominator).

0.321
0.136
__0.9457__
-0.056

See slides 20-23 of lecture on count outcomes

#lecture example - hits to Jeff Leek's website
```{r jleek example}
# this doesn't work yet!
download.file("https://dl.dropboxusercontent.com/u/7710864/data/gaData.rda",destfile="./data/gaData.rda")
load("./data/gaData.rda")
gaData$julian <- julian(gaData$date)
head(gaData)
```

```{r q4 - poisson on InsectSpray}
data(InsectSprays)
str(InsectSprays)
model1<-glm(count ~ spray,family="poisson",data=InsectSprays)
summary(model1)
```
##Question Five

# coefficients unchanged if time scale changed eg by factor 10

```{r q5 - log10 additional offset}
data(InsectSprays)
str(InsectSprays)
InsectSprays$t<-2
model1<-glm(count ~ spray,family="poisson",offset=log(t+1),data=InsectSprays)
summary(model1)
model2<-glm(count ~ spray,family="poisson",offset=log(t+1)+log(10),data=InsectSprays)
summary(model2)
```

##Question Six

Consider the data

x <- -5:5
y <- c(5.12, 3.93, 2.67, 1.87, 0.52, 0.08, 0.93, 2.05, 2.54, 3.87, 4.97)

Using a knot point at 0, fit a linear model that looks like a hockey stick with two lines meeting at x=0. Include an intercept term, x and the knot point term. What is the estimated slope of the line after 0?

2.037
-1.024
__1.013__
-0.183

```{r q6 knot point}
x <- -5:5
y <- c(5.12, 3.93, 2.67, 1.87, 0.52, 0.08, 0.93, 2.05, 2.54, 3.87, 4.97)
plot(x,y)
knot0=0
splineTerms <- sapply(knot0, function(knot) (x > knot) * (x - knot))
xMat <- cbind(1, x, splineTerms)
yhat <- predict(lm(y ~ xMat - 1))
plot(x, y, frame = FALSE, pch = 21, bg = "lightblue", cex = 2)
lines(x, yhat, col = "red", lwd = 2)
(yhat[11]-yhat[6])/(x[11]-x[6]) # =beta1 + beta2!
```

Copied code from Brian's example in Hodgepodge lecture 03_ 04_ a Fitting functions
slide 3
```{r example}
n <- 500; x <- seq(0, 4 * pi, length = n); y <- sin(x) + rnorm(n, sd = .3)
knots <- seq(0, 8 * pi, length = 20); 
splineTerms <- sapply(knots, function(knot) (x > knot) * (x - knot))
xMat <- cbind(1, x, splineTerms)
yhat <- predict(lm(y ~ xMat - 1))
plot(x, y, frame = FALSE, pch = 21, bg = "lightblue", cex = 2)
lines(x, yhat, col = "red", lwd = 2)
```