## Regression Models 

## Quiz One

### Question One

```{r Q1 manually}
library (manipulate)
library(dplyr)

x <- c(0.18, -1.54, 0.42, 0.95)
w <- c(2, 1, 3, 1)
mu<-c(rep(1.077,4),rep(0.1471,4),rep(0.0025,4),rep(0.300,4))
x<-c(rep(x,4))
w<-c(rep(w,4))
e<-w*(x-mu)^2
df<-data.frame(cbind(x,w,mu,e))
df
df %>% group_by(mu) %>% summarise(si=sum(e)) %>% arrange(si)
```

###Question Two

Consider the following data set

x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

Fit the regression through the origin and get the slope treating y as the outcome and x as the regressor.

(Hint, do not center the data since we want regression through the origin, not through the means of the data.)

```{r Q2}
x <- c(0.8, 0.47, 0.51, 0.73, 0.36, 0.58, 0.57, 0.85, 0.44, 0.42)
y <- c(1.39, 0.72, 1.55, 0.48, 1.19, -1.59, 1.23, -0.65, 1.49, 0.05)

df<-data.frame(cbind(x,y))

lm(y~x - 1)
```
```{r plot for Q2}
library(ggplot2)
g<-ggplot(df,aes(x=x,y=y))+
        geom_point()+
            geom_smooth(method=lm,   # Add linear regression line
                se=FALSE,formula=y~x-1) +
        xlim(0,1)
g
```


### Question Three

Do data(mtcars) from the datasets package and fit the regression model with mpg as the outcome and weight as the predictor. Give the slope coefficient.

-9.559
-5.344
0.5591
30.2851

```{r Q3}
data(mtcars)

lm(mpg~wt,data=mtcars)

```

Plot this

```{r plot for Q3}
library(ggplot2)
g<-ggplot(mtcars,aes(x=wt,y=mpg))+
        geom_point()+
            geom_smooth(method=lm,   # Add linear regression line
                se=FALSE,formula=y~x)+
        xlim(0,max(mtcars$wt))
g
```

### Question Six

```{r Q6}
x <- c(8.58, 10.46, 9.01, 9.64, 8.86)
z<-(x-mean(x))/sd(x)
z
```