---
title: "JHU_RM_Q3"
author: "mbh"
date: "Tuesday, June 16, 2015"
output: html_document
---

##Load and inspect data

```{r load data}
str(mtcars)
```
## Question One

Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor
variable and weight as confounder. 
Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.

-4.256
-3.206
-6.071
33.991

```{r q1}
fit<-lm(mpg~factor(cyl)+wt,data=mtcars)
summary(fit)
fitnf<-lm(mpg~cyl+wt,data=mtcars)
summary(fitnf)
```

## Question Three

Consider the mtcars data set. Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.InsectSprays

The P-value is small (less than 0.05). So, according to our criterion, we reject, which suggests that the interaction term is not necessary.
The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms is necessary.
The P-value is small (less than 0.05). Thus it is surely true that there is no interaction term in the true model.
The P-value is small (less than 0.05). So, according to our criterion, we reject, which suggests that the interaction term is necessary
The P-value is small (less than 0.05). Thus it is surely true that there is an interaction term in the true model.
The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms may not be necessary.

```{r q3}
fitcf<-lm(mpg~factor(cyl)+wt,data=mtcars)
fitint<-lm(mpg~factor(cyl)*wt,data=mtcars)

summary(fitcf)
summary(fitint)
```

##Question Four

```{r q4}
fit4<-lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
fit4
fit4nf<-lm(mpg ~ I(wt * 1) + factor(cyl), data = mtcars)
fit4nf
```

## Question Five

Question 5
Consider the following data set

x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
Give the hat diagonal for the most influential point

0.2025
0.9946
0.2804
0.2287

```{r q5}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
df<-data.frame(x,y)

fit5<-lm(y~x,data=df)
summary(fit5)
hatvalues(fit5)
```

## Question Six

```{r q6}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
df<-data.frame(x,y)
fit6<-lm(y~x,data=df)
dfbeta(fit6)
```